---
title: "MA615 Assignment 3"
author: "Andrew Zhang, Ningze Zu, Albert Ding, Jiahao Xu"
date: "9/30/2018"
output: pdf_document
---

https://www.kaggle.com/drgilermo/nba-players-stats#Seasons_Stats.csv

There is a schema for each of the variables available in the link above.

```{r, Andrew}
library(dplyr)
library(ggplot2)
season_ball <- read.csv("Seasons_Stats.csv")
season_ball$blank2 <- NULL
season_ball$blanl <- NULL
season_ball <- na.omit(season_ball)

## Looking and comparing win shares and number of three pointers made
## Comparing different eras of basketball to understand styles of play

## 2017 Season
ball_2017 <- season_ball %>% filter(Year == 2017)

ball_2017 %>% arrange(desc(X3P))
ball_2017 %>% arrange(desc(WS))

ggplot(ball_2017, aes(X3P, WS)) + geom_point() +
  geom_smooth(method = "lm") + 
  ggtitle("2017 3 Pointers vs. Win Shares")

## 2010 Season
ball_2010 <- season_ball %>% filter(Year == 2010)

ball_2010 %>% arrange(desc(X3P))
ball_2010 %>% arrange(desc(WS))

ggplot(ball_2010, aes(X3P, WS)) + geom_point() +
  geom_smooth(method = "lm") +
  ggtitle("2010 3 Pointers vs. Win Shares")

## 2003 Season
ball_2003 <- season_ball %>% filter(Year == 2003)

ball_2003 %>% arrange(desc(X3P))
ball_2003 %>% arrange(desc(WS))

ggplot(ball_2003, aes(X3P, WS)) + geom_point() +
  geom_smooth(method = "lm") + 
  ggtitle("2003 3 Pointers vs. Win Shares")

## Look at the number of 3 pointers made per team
## Notice the top teams from 2017 have some of the highest number of 3 pointers made

ball_2017 %>% group_by(Tm) %>% summarize(sum(X3P))
```

We can see a more positive trend for the 2017 data. This can possibly indicate more of a correlation between number of threes made and how well a team might do. 

In terms of 3 pointers vs win shares, it's hard to determine whether or not how well a team does is based on the number of 3 pointers a player might make. Although it seems like all players shoot 3 pointers in todays game, we need to explore this interaction further. 

```{r, age}
ball_2017 %>% arrange(desc(Age))

ggplot(ball_2017, aes(Age, MP)) + geom_point() +
  geom_smooth(method = "lm")

ggplot(ball_2010, aes(Age, MP)) + geom_point() +
  geom_smooth(method = "lm")

ggplot(ball_2003, aes(Age, MP)) + geom_point() +
  geom_smooth(method = "lm")
```

```{r, Albert}
library(dplyr)
library(ggplot2)
season_ball <- read.csv("Seasons_Stats.csv")
season_ball$blank2 <- NULL
season_ball$blanl <- NULL
season_ball <- na.omit(season_ball)

## Here, we will examine the trend of points per season between free throws and field goals. 

##Comparing different years of basketball will help us understand how offenses evolved in terms of scoring.  

##We have filtered only for years after 1980 because that was the year in which three pointers were first installed into the league, which we think may have affected the balance of offense. 

##Years after 1980:
threepointera <- season_ball %>% filter(Year >= 1980)


##Looking at overall field goal trends:
threepointera %>% arrange(desc(Year))

threepointera %>% group_by(Year) %>% summarize(sum(FG))

ggplot(threepointera, aes(Year, FG)) + geom_point() +
  geom_smooth(method = "lm")

##Looking at free throw trends:
threepointera %>% arrange(desc(Year))

threepointera %>% group_by(Year) %>% summarize(sum(FT))

ggplot(threepointera, aes(Year,FT)) + geom_point() +
  geom_smooth(method = "lm")

##Looking at field goal trends:
threepointera %>% arrange(desc(Year))

threepointera %>% group_by(Year) %>% summarize(sum(FG))

ggplot(threepointera, aes(Year,FG)) + geom_point() +
  geom_smooth(method = "lm")
```


```{r}
#Players' age over 25
ball1 <- filter(season_ball, !is.na(Age) & Age >= 25)
ball1

ball2 <- select(ball1, Year, Player, Tm, Age, Pos, G, PTS)
ball2

arrange(ball2, desc(PTS))

ggplot(ball2, aes(Age, PTS)) + geom_point() +
  geom_smooth(method = "lm")

#2008 Season
b1 <- filter(season_ball, Year==2008)
b2 <- select(b1, Year, Player, Tm, Age, Pos, G, X3P, X2P, PTS) 
b2

b2 %>%
  arrange(desc(PTS)) %>%
  summarise(total = n(), min.PTS = min(PTS), average.PTS = mean(PTS), max.PTS = max(PTS))

b2 %>%
  arrange(desc(X3P)) %>%
  summarise(total = n(), min.X3P = min(X3P), average.X3P = mean(X3P), max.X3P = max(X3P))

b2 %>%
  arrange(desc(X2P)) %>%
  summarise(total = n(), min.X2P = min(X2P), average.X2P = mean(X2P), max.X2P = max(X2P))
  
b2 %>%
  group_by(Tm) %>%
  summarise(N = n(), avg_PTS = mean(PTS, na.rm=TRUE)) 

ggplot(b2, aes(G,PTS)) + geom_point() + 
  geom_smooth(method = "lm")

ggplot(b2, aes(X3P,PTS)) + geom_point() + 
  geom_smooth(method = "lm")

ggplot(b2, aes(X2P,PTS)) + geom_point() + 
  geom_smooth(method = "lm")
```

```{r, Jiahao}
library(stringr)
season_ball <- read.csv("Seasons_Stats.csv")

# data in Season 2009 with players' age over 20 and made points are not zero
ball_2009_20 <- filter(season_ball,Age>20 & Year==2009 & !is.na(PTS))
ball1111 <- filter(ball_2009_20, !(str_detect(string = ball_2009_20 $Tm,pattern = "TOT"))) %>% group_by(Player) %>% 
  select(Year, Player, Pos, PER,Age, Tm, G,X3P, X2P, PTS ) %>%
  mutate(Average_PTS=PTS/G)  %>%
  arrange(desc(Average_PTS))
ball1111

# We want to figure out the relationship between Age and Average_PTS
ggplot(ball1111)+geom_point(mapping = aes(x=Age, y=Average_PTS))+geom_smooth(mapping = aes(x=Age, y=Average_PTS)) +ggtitle("Jiahao Xu's plot")
# We want to figure out the relationship between PER and Average_PTS, which is the scoring efficiency
ggplot(ball1111)+geom_point(mapping = aes(x=PER, y=Average_PTS))+geom_smooth(mapping = aes(x=PER, y=Average_PTS)) +ggtitle("Jiahao Xu's plot")
# We want to figure out the relationship between 2 points made and PER
ggplot(ball1111)+geom_point(mapping = aes(x=X2P, y=PER))+geom_smooth(mapping = aes(x=X2P, y=PER)) +ggtitle("Jiahao Xu's plot")
```

Conclusion
